Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-04 WSSSPE2_report/WSSSPE2_review.tex
Front matter and sections 1-2: tiny changes. Section 3: re-organizing...to be cont'd.

\begin{abstract} \todo{TBD due date above}       
\todo{update abstract}
This technical report discusses the submission and peer-review process used by
the Second Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE2) and the results of that process. It is intended to record
both the alternative submission and program organization model used by WSSSPE2
as well as the papers associated with the workshop that resulted from that
process.
\end{abstract}

@@

31 submissions were received, and all but one used either
arXiv\footnote{\url{http://arxiv.org}} or
figshare\footnote{\url{http://figshare.com}} to self-publish their papers.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keynotes} \label{sec:keynotes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Lead: Choi. Volunteers welcome---contact lead}

\todo{Live notes at \href{http://tinyurl.com/q6ew45v}{Google Notes}.
%https://docs.google.com/document/d/1-BxkYWDQ6nNNBXBStUL0xcKF9qCTlEALwf928J_MemI/edit?usp=sharing
Abstracts at \href{http://tinyurl.com/mnenzms}{WSSSPE2}.}

The workshop featured two keynote addresses. In the opening keynote
presentation, Kaitlin Thaney of Mozilla Science talked about her organization's
work and policy to enable and support sustainable and reproducible scientific
research. Our second keynote speaker was Neil Chue Hong of Software
Sustainability Institute. He shined the light on how scientific software is
prevalently driving advances in many sciences and engineering fields. Following
the keynotes, participants broke into groups to define and discuss the crucial
notion ``sustainability'' spawned by the speeches of our keynote speakers as well
as the theme of our workshop.

@@

\todo{Slides: \url{http://tinyurl.com/las8tsq}}

\begin{enumerate}
\item Why is Mozilla interested in Science: Mozilla is a non-profit interested
in openness, in news, in site creation, and in Science, taking advantage of the
open web.
\item Current systems designed to create friction --- social issues hindering us
\item case example: National Cancer Institute's caBIG
\begin{itemize}
\item $350m$ spent, $60m+$ for management, $70 + $ tools created, but still seen
as a failure (Kaitlin didn't really say why? Were the tools used? Did they
immediately bit-rot? Was their development unsustained?). - should ask question
about this
\item Those that had the least investment were the most used, the largest
investments were the least used.
\item Is there a caBIG report available? Quotes from Andrea Califano/Joe Gray,
looks like there is:
\url{http://tinyurl.com/maf6dz2}%http://www.informationweek.com/architecture/report-blasts-problem-plagued-cancer-research-grid/d/d-id/1097068?
\end{itemize}

\item What do we mean by open research? In regard to
community/technology/practices
\item Inefficiency: \url{http://www.xkcd.com/1445/}
\item Need research tools (e.g. software repositories), social capital (e.g.
incentives) and capacity (eg training and mentorship) for efficient reproducible
research
\item Our systems need to talk to each other but Q for speaker is that as
systems become less monolithic, paradoxically it may be harder to sustain the
links between them see:
\begin{itemize}
\item
\url{http://tinyurl.com/l76tba2}%http://www.slideshare.net/jameshowison/scientific-software-sustainability-and-ecosystem-complexity
(but does Anon Grizzly have other cites/info for that?)
\item Note, for example, that creative commons licensed works typically don't
have dependencies (a book, a photo, an artwork).
\end{itemize}

\item Software Discovery Index: \url{http://softwarediscoveryindex.org/report/}
\item Final points:
\begin{itemize}
\item Design with the audience not the individual in mind
\item Design to unlock latent potential of our systems
\item Rethink how we reward researchers and support roles
\item By mindful of jargon/semantics traps
\end{itemize}
\item Questions:
\begin{itemize}
\item Ted Habermann: if the metadata is minimal less onerous for data provider,
but more onerous for user. How to work with data providers to show that users
need more.
\item Find balance between things that have to be captured and what would be
ideal. Don't want to get to point where like old Harvard Dataverse, three pages
of metadata capture, only first four things ever filled out. Don't want to lose
engagement.
\end{itemize}
\end{enumerate}

>>>>

\begin{abstract}      
\todo{update abstract}
This technical report discusses the submission and peer-review process used by
the Second Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE2) and the results of that process. It is intended to record
both the alternative submission and program organization model used by WSSSPE2
as well as the papers associated with the workshop that resulted from that
process.
\end{abstract}

@@

31 submissions were received; all but one used  
arXiv\footnote{\url{http://arxiv.org}} or
figshare\footnote{\url{http://figshare.com}} to self-publish their papers.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keynotes} \label{sec:keynotes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{\footnotesize{Lead: Choi. Volunteers contact lead. Live notes: \href{http://tinyurl.com/q6ew45v}{Google Notes}}.
%https://docs.google.com/document/d/1-BxkYWDQ6nNNBXBStUL0xcKF9qCTlEALwf928J_MemI/edit?usp=sharing
Abstracts: \href{http://tinyurl.com/mnenzms}{WSSSPE2}. Slides: \href{http://tinyurl.com/las8tsq}{here}.}

The workshop featured two keynote addresses. In the opening keynote
presentation, Kaitlin Thaney of the Mozilla Science Lab talked about her
organization's work and policy to enable and support sustainable and
reproducible scientific research. Our second keynote speaker was Neil Chue Hong
of Software Sustainability Institute. He shined the light on how scientific
software is prevalently driving advances in many science and engineering fields.
Both keynote speeches spawned further discussion among workshop participants on
the crucial notion ``sustainability'' in the theme of our workshop.

@@

Kaitlin Thaney is the Director of the Mozilla Science Lab (hereafter Mozilla).
Mozilla is a non-profit organization interested in openness, in news, in website
creation, and in Science, taking advantage of the open web.

Thaney started noting the unfortunate fact that many current systems have been designed to create friction hindering us, despite the original intentions.%---social issues hindering us
An example is the National Cancer Institute's caBIG. 
A total of $\$350m$ was spent, more than $\$60m$ for management, $70+$ tools created, but still seen
as a failure (Kaitlin didn't really say why? Were the tools used? Did they
immediately bit-rot? Was their development unsustained?)-should ask question
about this.
 Those that had the least investment were the most used, the largest
investments were the least used.
 Is there a caBIG report available? Quotes from Andrea Califano/Joe Gray,
looks like there is:
\url{http://tinyurl.com/maf6dz2}.%http://www.informationweek.com/architecture/report-blasts-problem-plagued-cancer-research-grid/d/d-id/1097068?

What do we mean by open research? In regard to
community/technology/practices

Inefficiency: \url{http://www.xkcd.com/1445/}. Need research tools (e.g. software repositories), social capital (e.g.
incentives) and capacity (eg training and mentorship) for efficient reproducible
research.

Our systems need to talk to each other but Q for speaker is that as
systems become less monolithic, paradoxically it may be harder to sustain the
links between them see:
\url{http://tinyurl.com/l76tba2}%http://www.slideshare.net/jameshowison/scientific-software-sustainability-and-ecosystem-complexity
(but does Anon Grizzly have other cites/info for that?).
Note, for example, that creative commons licensed works typically do not 
have dependencies (a book, a photo, an artwork).
Software Discovery Index: \url{http://softwarediscoveryindex.org/report/}.

 Final points:
\begin{itemize}
\item Design with the audience not the individual in mind
\item Design to unlock latent potential of our systems
\item Rethink how we reward researchers and support roles
\item By mindful of jargon/semantics traps
\end{itemize}

Questions:
\begin{itemize}
\item Ted Habermann: if the metadata is minimal less onerous for data provider,
but more onerous for user. How to work with data providers to show that users
need more.
\item Find balance between things that have to be captured and what would be
ideal. Don't want to get to point where like old Harvard Dataverse, three pages
of metadata capture, only first four things ever filled out. Don't want to lose
engagement.
\end{itemize}
